\subsection{Expert Advice}\label{expertadvie}

Over the last few months contact has been made with various experts in their fields which are related to this project. These people are as follows:

\begin{itemize}

\item Dr Mark Wright

Involved with a group which is working on a very similar project, Dr Mark Wright is an extremely helpful source of information about how to solve implementation problems. His groups goals are slightly different in that they are interested in making this data available, however, with their partnership with Lothian Buses and the time scale they are aiming for, I am hopeful that we can work together to achieve both of our goals. 

\item Dr John Moncrieff 

A professor of Micrometeorology at the \emph{University of Edinburgh}, Dr John Moncrieff, has a research interest in gas diffusion modelling in the earth's atmosphere. With his guidance I am developing a key insight into the optimal model to use to interpolate the sparse data sets generated. 

\item Dr Stuart Sneddon

As the business manager at \emph{Air and Environment (Scotland)}, Dr Sneddon has kindly agreed to meet with me in the coming month to discuss his companies atmospheric readings. These readings can be used to calibrate the sensors and ensure that their readings are accurate. The data can also be used for training and testing models. 

\end{itemize}

I have been in contact with various other people briefly, but wish to thank in particular the above people for their help in gathering information for this project.
